**Microbiota predict _Clostridium difficile_ severity in germ-free mice colonized with human feces**

### 2016-12-07 submission/Manuscript.Rmd

Gnotobiotic or germ-free mouse models have been used for a range of studies of CDI, including assessment of species-specific interactions between _C. difficile_ and competing microbial community members [@Reeves2012], analysis of nutrient restriction [@Ng2013], _in vivo_ transcriptomics of _C. difficile_ and examination of host immune response to CDI [@Pawlowski2012]. Further, CDI therapeutics such as antibiotics and fecal microbiota transplants have been tested extensively in a gnotobiotic-piglet or piglet-to-gnotobiotic-mouse model of disease [@Kim2014;@Diao2016]. Pigs have a longer digestive tract with components more similar to humans than mice and are typically infected by strains typical in human infection [@Steele2010]. However, the murine and porcine microbiomes typically do not resemble those of the human gut.

@@

Here we colonized germ-free mice with a variety of human stool communities and allowed these communities to colonize the mice for 2 weeks, prior to _C. difficile_ challenge. It is important to note that the mouse gut environment is different from that of the human, and as such exerts a selective pressure on the human stool microbial community that stabilizes over the 2 weeks. Thus these mice cannot be considered "humanized" but rather colonized by human-associated microbes. Interestingly, clinical status was not associated with proclivity to severe _C. difficile_ infection as some healthy donor mice were susceptible. A surprising finding was that none of the human-associated mouse communities were completely resistant to colonization by _C. difficile_. This is in contrast to numerous reports of mouse microbiome studies [@Schubert2015] and the knowledge that in general healthy humans without gut perturbations are resistant to _C. difficile_ infection [Heinlen2010]. On the other hand, it is known that humans, especially health care employees, can be asymptomatic carriers of _C. difficile_ [@FuruyaKanamori2015]. This may explain the fact that the majority of our groups of mice were colonized by _C. difficile_ but did not display outward signs of illness or infection-associated weight loss (Fig 2).

@@

Further experiments will test if infecting susceptible communities with weak strains allows the mice to survive. It has been recently reported that carriage of a non-toxigenic strain of _C. difficile_ can prevent additional colonization or recurrence by a pathogenic strain, suggesting a role of niche-filling or immunity this pre-colonization provides [@Gerding2015]. One interesting experiment would be to recapitulate this finding in human-associated mouse microbiomes by making a susceptible community resistant to infection. Further work with human-associated communities in mice and mathematical modeling will allow for uncovering of complex human-associated disease processes and dynamics as they relate to human _C. difficile_ infection.

@@

**Human samples** Donor stool samples were chosen from a sample bank collected from patients from October 2010 to November 2012 at the University of Michigan. Sample collection methods and study was approved by the Institutional Review Board at the University of Michigan. Stool samples from University of Michigan Health System patients were submitted to the microbiology lab for _C. difficile_ testing. _C. difficile_ presence was measured using C.Diff Quik Chek Complete Assay (Techlab). Samples from these patients were stored in Cary-Blair medium at -80C. All patients consented to participation in the study. Donor stool for this study was selected from one patient that tested positive for _C. difficile_ infection and five patients that presented with diarrhea and tested negative for _C. difficile_. Healthy donor stool was collected from adult volunteers living in the Ann Arbor, MI area. 100ul of stool slurry was gavaged into each mouse.

@@

##Acknowledgments## 
We would like to thank members of the Schloss lab and Jhansi Leslie for analysis advice. We would also like to thank the Microbial Systems Laboratory at the University of Michigan for sequencing assistance.

@@

**Figure 5. Infection of mice with different _C. difficile strains_.** 3 strains of _C. difficile_ were used to infect mice colonized with susceptible (DA00578) or resistant (DA00369, DA00430) human donor stool. A) _C. difficile_ stool CFU was enumerated over 10 days. B) Percent weight loss was calculated each day for each mouse. In both plots, each mouse is a point and lines represent the mean of each cage.

>>>>

Gnotobiotic or germ-free mouse models have been used for a range of studies of CDI, including assessment of species-specific interactions between _C. difficile_ and competing microbial community members [@Reeves2012], analysis of nutrient restriction [@Ng2013], _in vivo_ transcriptomics of _C. difficile_ and examination of host immune response to CDI [@Pawlowski2010]. Further, CDI therapeutics such as antibiotics and fecal microbiota transplants have been tested extensively in a gnotobiotic-piglet or piglet-to-gnotobiotic-mouse model of disease [@Kim2014;@Diao2016]. Pigs have a longer digestive tract with components more similar to humans than mice and are typically infected by strains typical in human infection [@Steele2010]. However, the murine and porcine microbiomes typically do not resemble those of the human gut.

@@

Here we colonized germ-free mice with a variety of human stool communities and allowed these communities to colonize the mice for 2 weeks, prior to _C. difficile_ challenge. It is important to note that the mouse gut environment is different from that of the human, and as such exerts a selective pressure on the human stool microbial community that stabilizes over the 2 weeks. Thus these mice cannot be considered "humanized" but rather colonized by human-associated microbes. Interestingly, clinical status was not associated with proclivity to severe _C. difficile_ infection as some healthy donor mice were susceptible. A surprising finding was that none of the human-associated mouse communities were completely resistant to colonization by _C. difficile_. This is in contrast to numerous reports of mouse microbiome studies [@Schubert2015] and the knowledge that in general healthy humans without gut perturbations are resistant to _C. difficile_ infection [@Heinlen2010]. On the other hand, it is known that humans, especially health care employees, can be asymptomatic carriers of _C. difficile_ [@FuruyaKanamori2015]. This may explain the fact that the majority of our groups of mice were colonized by _C. difficile_ but did not display outward signs of illness or infection-associated weight loss (Fig 2).

@@

It has been recently reported that carriage of a non-toxigenic strain of _C. difficile_ can prevent additional colonization or recurrence by a pathogenic strain, suggesting a role of niche-filling or immunity this pre-colonization provides [@Gerding2015]. One interesting experiment would be to recapitulate this finding in human-associated mouse microbiomes by making a susceptible community resistant to infection. Further work with human-associated communities in mice and mathematical modeling will allow for uncovering of complex human-associated disease processes and dynamics as they relate to human _C. difficile_ infection.

@@

**Human samples** Donor stool samples were chosen from a sample bank collected from patients from October 2010 to November 2012 at the University of Michigan Health System. Sample collection methods and the study were approved by the Institutional Review Board at the University of Michigan. Stool samples from University of Michigan Health System patients were submitted to the microbiology lab for _C. difficile_ testing. _C. difficile_ presence was measured using C.Diff Quik Chek Complete Assay (Techlab). Fecal samples from these patients were stored in Cary-Blair medium at -80C. All patients consented to participation in the study. Donor stool was selected from one patient that tested positive for _C. difficile_ infection and five patients that presented with diarrhea and tested negative for _C. difficile_. Healthy donor stool was collected from adult volunteers living in the Ann Arbor, MI area. 100ul of stool slurry was gavaged into each mouse.

@@

##Acknowledgments## 
We would like to thank members of the Schloss lab and Jhansi Leslie for experimental and analytical advice. We would also like to thank the Microbial Systems Laboratory at the University of Michigan for sequencing assistance.

@@
